---
title: "Ecom Store"
author: "Prashidha Shrestha"
date: "2022/07/07"
output:
  html_document:
    code_folding: show
  pdf_document: default
  word_document:
    topic: yes
    always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clearing the working space and setting a path.

```{r}
# remove(list=ls())
### Setting working directory. 
## setwd("D:/DataDoorGit/R Data Science Projects")
```

## Import all necessary libraries.

```{r message=FALSE, warning=FALSE}

library(ggplot2) 
library(skimr) 
library(dplyr) 
library(scales) 
library(reshape)
library(tidyr)
library(magrittr)
library(corrplot)
```


## Read and Import csv file and store it in a variable. 

```{r}
dframe <- read.csv("ecom_store.csv") # Read the csv file and create a new data frame.
```


## Check the dimension of the whole dataframe.

```{r}
dim(dframe) # Displays the dimesnion of the dataframe.
```
```{r}
head(dframe) # It displays some of the top rows.
```

## Check the last few rows of the dataframe.

```{r}
tail(dframe) # It displays some of the last rows.
```

## Check the dataframe by using str of dataframe

```{r}
str(dframe) # understand dataframe and its datatypes with their values.
```

## Use of sum to check the total rows with na value in dataframes.

```{r}
# first check empty or missing values in overall dataframe. 
sum(is.na(dframe)) # It shows total number of missing values.
```

## Assign a variable to chcck na values for each columns.

```{r}
# check whether columns have either NA or empty cells.
empty_columns <- colSums(is.na(dframe) | dframe == "" | dframe == " ")
empty_columns
```

## Check the total length of postal code column with na values.

```{r}
# check empty rows from postalcode column Which has almost 40k rows 
dframe_isna <- dframe[(is.na(dframe$Postal.Code) | dframe$Postal.Code=="" | dframe$Postal.Code==" "), ]
length(dframe_isna$Postal.Code)
```

## Remove unnecessary features from the dataframe.
 
```{r}
## Remove RowID and Country as Country has only one value in all rows 'US'. 
dframe = subset(dframe, select = -c(Country,Row.ID,Postal.Code))
```
 
## Calcuate Shipping time on the basis of SHip date and Order Date.  
 
```{r}
## Creating new column  Shipping Time
dframe$shipping_time <- as.Date(as.character(dframe$Ship.Date),
    format="%Y-%m-%d") - as.Date(as.character(dframe$Order.Date), format="%Y-%m-%d")
```

## Create columns month, day and year from Order Date.

```{r}
dframe$order<-dframe$Order.Date
dframe<-separate(dframe,order,c("month","day","year"),sep="-")
```

## we see that there is an outlier in the Sales feature, an unusual hike. We should replace it with the mean of sales.

```{r}
maxSales <- max(dframe$Sales)
dframe$Sales <- replace(dframe$Sales, dframe$Sales==maxSales,mean(dframe$Sales))
```

## Check the summary of the dataframe.

```{r}
summary(dframe)
```

## Use sskimr library to understand the dataframe in simple steps from different perspective.

```{r}
skim(dframe)
```

## Bar graph for counting the total number of orders as per the Markets

```{r}
region_order <- table(dframe$Market)
barplot(region_order, main="Total Orders by Markets",xlab="Markets",col="lightblue")
```

## Distribution of Quantity.

```{r}
hist(dframe$Quantity, main="Frequency Distribution of Quantity",xlab="Quantity Ordered",ylab="Frequency",col="lightpink")
```

## Pie chart Visualizations For Category as per Total Sales data.

```{r}
gd <- dframe %>% group_by(Category) %>% summarize(Sales=sum(Sales))
pct <- round(gd$Sales/sum(gd$Sales)*100)
lbls <- paste(gd$Category,pct)
lbls <- paste(lbls,"%", sep = " ")
colors = c('lightskyblue','plum2','peachpuff')
pie(gd$Sales, labels=lbls, main="", col=colors)
```

## Filter the numerical features for correlation plot.

```{r}
## Subset specific features for plotting correlations between.
num_data_cols <- dframe[,c(16,17,18,19,20)]
data.cor = cor(num_data_cols)
data.cor
```

```{r}
corrplot(cor(data.cor), method = "number",type = "upper")
```

## Scatter plot to display the Sales and profit as Ship Mode.


```{r}
ggplot(data = dframe, aes(x = Sales, y = Profit, color = Ship.Mode)) + geom_point()
```

## Scatter plot to display the Sales and Profit relation as Category 

```{r}
ggplot() + geom_point(data = dframe, aes(x = Sales, y = Profit, color = Category))
```

## Bar plots to display sub category and their profit as per market. 

```{r}
ggplot() + geom_bar(data = dframe, aes(x = Sub.Category, y = Profit, fill = Market), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Bar plot to display the Sales of category as Order Priority.

```{r}
ggplot() + geom_bar(data = dframe, aes(x = Category, y = Sales, fill = Order.Priority), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Bar plot to display the Market place and number of orders with respect to segments.

```{r}
ggplot(dframe,aes(y=Market,fill=Segment)) + 
  geom_bar(position = "stack") + labs(title = 'Total number of Orders based on Segments and Market Areas.',y='Market Place',x='Total number of Orders') + facet_wrap(~Segment)
```

## Use of facetwrap to display the Ordered Catgeory with respect to Order Priorities.

```{r}
ggplot(dframe,aes(x=Category,fill=Order.Priority)) + 
  geom_bar(position = "stack") + labs(title = 'Category',
                                      x='Categories',y='Total number of Orders')+ facet_wrap(~Order.Priority)
```

## Subset data for visualizing plots

```{r}
## Group by Order priority and calculate their mean Shipping COsts.
df_ship_cost__mean <- aggregate(dframe$Shipping.Cost, list(dframe$Order.Priority), FUN=mean)
names(df_ship_cost__mean)[1]<- paste("Priorities")
names(df_ship_cost__mean)[2]<- paste("Mean")
df_ship_cost__mean
```


```{r}
ggplot(df_ship_cost__mean,aes(x=Priorities, y = Mean, fill=Priorities))+
  geom_bar( stat = "identity")+
  labs(title = 'Mean Shipping cost plot of Orders Grouped by Priorities',x='Priorities',y='Mean Value')
```



```{r}
## Select specific ranges of numeric columns 
num_cols <-dframe %>% dplyr::select(where(is.numeric))
num_cols<-num_cols[,1:5]
```

## Test Pearsons correlation between Sales and Quantity.

```{r}
res1 <- cor.test(dframe$Sales, dframe$Quantity, 
                method = "pearson")
res1
```

## Test Pearsons correlation between Sales and Discount

```{r}
res2 <- cor.test(dframe$Sales, dframe$Discount, 
                method = "pearson")
res2
```

## Test Pearsons correlation between Sales and Profit

```{r}
res3 <- cor.test(dframe$Sales, dframe$Profit, 
                method = "pearson")
res3
```

## Test Pearsons correlation between Sales and Shipping Cost

```{r}
res4 <- cor.test(dframe$Sales, dframe$Shipping.Cost, 
                method = "pearson")
res4
```


```{r}
df_standardize <- as.data.frame(scale(num_cols))
data_sorted_stand <- df_standardize[order(df_standardize$Sales, decreasing = TRUE), ]
after_sort_stand = data_sorted_stand[-c(1:20),]
boxplot(after_sort_stand)
```


```{r}
data_sorted_stand_profit <- after_sort_stand[order(after_sort_stand$Profit,decreasing = TRUE), ]
after_sort_stand_v2 = data_sorted_stand_profit[-c(1:5),]
dframe<-after_sort_stand_v2
boxplot(after_sort_stand_v2)
```

```{r}
data_sorted_stand2 <- data_sorted_stand_profit[order(data_sorted_stand_profit$Profit), ]
after_sort_stand2 = data_sorted_stand2[-c(1:5),]
dframe<-after_sort_stand2
boxplot(after_sort_stand2)
```


```{r}
bc <- select(dframe,Sales,Profit,Shipping.Cost,Quantity,Discount)
bc_model_full <- lm(Sales ~ Shipping.Cost + Profit + Quantity + Discount, data=dframe)
bc_model_full
```

```{r}
preds <- predict(bc_model_full)

plot(preds, bc$Sales, xlab = "Prediction", ylab = "Observed")
abline(a = 0, b = 1)
```

```{r}
summary(bc_model_full)
```


```{r}
set.seed(123)
ind <- sample(2, nrow(bc), replace=TRUE, prob=c(0.80, 0.20))
bc_train <- bc[ind==1,]
bc_test <- bc[ind==2,]


#Let's build now a linear regression model using the training data and print it:
(bc_model <- lm(Sales ~ Shipping.Cost + Profit + Quantity+Discount, data=bc_train))

#We can also view the model's summary
summary(bc_model)


######Evaluating graphically
#Let's make predictions on our training dataset and store the predictions as a new column
bc_train$pred <- predict(bc_model)

# plot the ground truths vs predictions for training set
ggplot(bc_train, aes(x = pred, y = Sales)) +
  geom_point() +
  geom_abline(color = "blue")
```


```{r}
bc_test$pred <- predict(bc_model , newdata=bc_test)
```

```{r}
# plot the ground truths vs predictions for test set and examine the plot. Does it look as good with the predictions on the training set?
ggplot(bc_test, aes(x = pred, y = Sales)) +
  geom_point() +
  geom_abline(color = "blue")
```

```{r}
bc_mean <- mean(bc_test$Sales)
tss <- sum((bc_test$Sales - bc_mean)^2)
err <- bc_test$Sales-bc_test$pred
rss <- sum(err^2)
(rsq <- 1-(rss/tss))
```

```{r}
#Calculate residuals
res <- bc_train$Sales-bc_train$pred
#For training data we can also obtain the residuals using the bc_model$residuals

# Calculate RMSE, assign it to the variable rmse and print it
(rmse <- sqrt(mean(res^2)))


# Calculate the standard deviation of actual outcome and print it
(sd_bc_train <- sd(bc_train$Sales))
```







